grad_weight

### 2022-03-30 inst/surfing_jsa_rebuttal.tex

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

\section*{MS596 ``Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods": rebuttal}

Below, we reproduce the two reviewers' comments; our replies to the
issues are in \textcolor{blue}{blue}.  We have indicated changes to
the manuscript where appropriate.

### 2022-04-12 inst/grad_weight.Rmd
more examples (there is a bug somewhere)

## A slightly more complicated example

Two chess players, 1, and 2, play chess.  Our observations (chess
games, say) comprise $a+b+c$ games in which $p_1$ played White,
winning $a$ games, drawing $b$ and losing $c$ [we write $+a=b-c$]; and
$d+e+f$ games in which $p_2$ played White, with $+d=e-f$.  Consider
the following likelihood function:

$$
\frac{(w p_1)^a\cdot(D(p_1+p_2))^b\cdot p_2^c}{(w p_1+D(p_1+p_2)+p_2)^{a+b+c}}
\frac{(w p_2)^d\cdot(D(p_1+p_2))^e\cdot p_1^f}{(w p_2+D(p_1+p_2)+p_1)^{d+e+f}}
$$

@@

brackfunc <- function(w,D){
  list(
  b1 = c(p1=w),
  b2 = c(p1=D, p2=D),	
  b3 = c(p2=1),
  b4 = c(p1=w+D,p2=1+D),
  b5 = c(p2=w),
  b6 = c(p1=1),
  b7 = c(p1=1+D,p2=w+D)
)
}
brackfunc(w,D)

wanted <- list(NULL,c("p1","p2"),NULL,c("p1","p2"), NULL,NULL,c("p1","p2"))
wanted

`f2` <- function(w,D, a=3,b=12,c=1,d=5,e=6,f=6){
  H <- hyper3()	      
  brackets <- brackfunc(w,D)
  H[brackets[[1]]] %<>% inc(a)	
  H[brackets[[2]]] %<>% inc(b)	
  H[brackets[[3]]] %<>% inc(c)	
  H[brackets[[4]]] %<>% dec(a+b+c)
  H[brackets[[5]]] %<>% inc(d)	
  H[brackets[[2]]] %<>% inc(e)	 # sic, the draw brackets are the same
  H[brackets[[6]]] %<>% inc(f)	
  H[brackets[[7]]] %<>% dec(d+e+f)
  return(H)
}

@@

$$
\frac{(w p_1)^3\cdot(D(p_1+p_2))^{12}\cdot p_2^1}{(w p_1+D(p_1+p_2)+p_2)^{16}}
\frac{(w p_2)^5\cdot(D(p_1+p_2))^6\cdot p_1^6}{(w p_2+D(p_1+p_2)+p_1)^{17}}
$$

@@

```{r}
diff <- 0
dw <- 1e-4
probs <- c(p1=0.35 , p2=0.65)
B <- brackfunc(w,D)
for(i in seq_along(wanted)){
                                   diff <- diff +  grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]])
                         print(grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]]))

>>>>

# Slightly more complicated example

```{r}
H <- hyper3()
f <- function(w){
  H <- hyper3()
  H[c(p1=1 , p2=w , p3=w , p4=1)] <- 5
  H
}
w <- 1.4
dw <- 1e-6
HH <- f(w)
HH
probs <- c(p1=0.1,p2=0.2,p3=0.3,p4=0.4)
wpv <- c(p1=1 , p2=w , p3=w , p4=1)
LHS <- grad_weight_single(HH, probs, wpv, c("p2","p3"))
RHS <-  (loglik(probs,f(w=w+dw)) - loglik(probs,f(w=w)))/dw
c(RHS,LHS,LHS-RHS)
```

## An example from chess

Two chess players, 1, 2, and 3, play chess.  Our observations (chess
games, say) comprise $a+b+c$ games in which $p_1$ played White,
winning $a$ games, drawing $b$ and losing $c$ [we write $+a=b-c$]; and
$d+e+f$ games in which $p_2$ played White, with $+d=e-f$.  There is
also a single game between $p_3$ (white) who beat $p_1$ (black).
Consider the following likelihood function:

$$
\frac{(w p_1)^a\cdot(D(p_1+p_2))^b\cdot p_2^c}{(w p_1+D(p_1+p_2)+p_2)^{a+b+c}}
\frac{(w p_2)^d\cdot(D(p_1+p_2))^e\cdot p_1^f}{(w p_2+D(p_1+p_2)+p_1)^{d+e+f}}
\frac{ w p_3}{wp_3+p_1}
$$

@@

brackfunc <- function(w,D){
  list(
  b1 = c(p1=w),
  b2 = c(p1=D, p2=D),	
  b3 = c(p2=1),
  b4 = c(p1=w+D,p2=1+D),
  b5 = c(p2=w),
  b6 = c(p1=1),
  b7 = c(p1=1+D,p2=w+D),
  b8 = c(p3=w),
  b9 = c(p3=w,p1=1)
)
}
brackfunc(w,D)

wanted <- list(NULL,c("p1","p2"),NULL,c("p1","p2"), NULL,NULL,c("p1","p2"),NULL,NULL)
wanted

`f2` <- function(w,D, a=3,b=12,c=1,d=5,e=6,f=6){
  H <- hyper3()	      
  brackets <- brackfunc(w,D)
  H[brackets[[1]]] %<>% inc(a)	
  H[brackets[[2]]] %<>% inc(b)	
  H[brackets[[3]]] %<>% inc(c)	
  H[brackets[[4]]] %<>% dec(a+b+c)
  H[brackets[[5]]] %<>% inc(d)	
  H[brackets[[2]]] %<>% inc(e)	 # sic, the draw brackets are the same
  H[brackets[[6]]] %<>% inc(f)	
  H[brackets[[7]]] %<>% dec(d+e+f)
  H[brackets[[8]]] %<>% inc
  H[brackets[[9]]] %<>% dec
  return(H)
}

@@

$$
\frac{(w p_1)^a\cdot(D(p_1+p_2))^b\cdot p_2^c}{(w p_1+D(p_1+p_2)+p_2)^{a+b+c}}
\frac{(w p_2)^d\cdot(D(p_1+p_2))^e\cdot p_1^f}{(w p_2+D(p_1+p_2)+p_1)^{d+e+f}}
\frac{ w p_3}{wp_3+p_1}
$$

@@

```{r}
diff <- 0
dw <- 1e-4
probs <- c(p1=0.15 , p2=0.45, p3=0.4)
B <- brackfunc(w,D)
for(i in seq_along(wanted)){
                                   diff <- diff +  grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]])
                         print(grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]]))
